Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
Rearrange ggplot layers for power cmp

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                select(-(RESET5_p_value:RESET10_p_value)) %>%
                select(-RESET3_p_value, -F_p_value) %>%
                rename(RESET_p_value = RESET4_p_value) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(RESET_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  geom_point(data = ~.x %>% 
               group_by(log_effect_size, reject_dirichlet) %>% 
               summarise(num_response = n()),
             aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = -Inf, yend = Inf), alpha = 0.3, col = "grey") +
  theme_light(base_size = 5) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_size_manual(values = c(1, 2, 3)) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  xlab("Effect size (log e)") +
  ylab("Power") +
  labs(col = "", size = "#lineups")

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                select(-(RESET5_p_value:RESET10_p_value)) %>%
                select(-RESET3_p_value, -F_p_value) %>%
                rename(RESET_p_value = RESET4_p_value) %>%
                pivot_longer(RESET_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  geom_point(data = ~.x %>% 
               group_by(log_effect_size, reject_dirichlet) %>% 
               summarise(num_response = n()),
             aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = -Inf, yend = Inf), alpha = 0.3, col = "grey") +
  theme_light(base_size = 5) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  scale_size_manual(values = c(1, 2, 3)) +
  xlab("Effect size (log e)") +
  ylab("Power") +
  labs(col = "", size = "#lineups")

>>>>

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_point(data = ~.x %>% 
             group_by(log_effect_size, reject) %>% 
             summarise(num_response = n()),
           aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3)  +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                select(-(RESET5_p_value:RESET10_p_value)) %>%
                select(-RESET3_p_value, -F_p_value) %>%
                rename(RESET_p_value = RESET4_p_value) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(RESET_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject, times = 500), 
              aes(log_effect_size, reject, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_size_manual(values = c(1, 2, 3)) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  xlab("Effect size (log e)") +
  ylab("Power") +
  labs(col = "", size = "# lineups")

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_point(data = ~.x %>% 
               group_by(log_effect_size, reject) %>% 
               summarise(num_response = n()),
             aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3) +
  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                select(-(RESET5_p_value:RESET10_p_value)) %>%
                select(-RESET3_p_value, -F_p_value) %>%
                rename(RESET_p_value = RESET4_p_value) %>%
                pivot_longer(RESET_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject, times = 500), 
              aes(log_effect_size, reject, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  scale_size_manual(values = c(1, 2, 3)) +
  xlab("Effect size (log e)") +
  ylab("Power") +
  labs(col = "", size = "# lineups")

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.tex
title change
Rearrange ggplot layers for power cmp

\title{Why aren't significance tests commonly used for linear regression
diagnostics?}

@@

\citet{vanderplas2021statistical} suggest that \(\alpha\) can be
estimated using MLE with Equation \ref{eq:c-interesting-expectation}.
But for precise estimate of \(\alpha\), additional responses to
Rorschach lineups, which is a type of lineup consists of only null
plots, are required.

@@

@@

As mentioned in Section \ref{calculating-p-values}, \(\alpha\) used in
Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using
Rorschach lineups. Hence, 36 lineups consist of only null plots with all
combinations of \(n\) and \(X_{raw}\) and three replications are
designed to be included in experiment III. Note that we generated the
lineups using the same simulation mechanism discussed in Section
\ref{sampling-from-the-null-distribution} rather than from a null model.
We overlooked this detail until the data had been collected. Thus, we
used another valid method suggested in vanderplas2021statistical which
was designed for typical lineups to estimate \(\alpha\) correctly.
However, by not using Rorschach lineups, we did lose some accuracy in
the estimation. We recommend future studies to generate Rorschach
lineups strictly following the definition given by
buja\_statistical\_2009.

@@

@@

For the non-linearity model, the power of all four tests increases as
the effect gets larger. The power curve of F-test climbs aggressively
from 25\% to around 90\% as \(log_e(E)\) increases from 0 to 2, while
power of other tests respond inactively to the change of effect and
remain lower than 25\% throughout the period, showing that F-test is way
more sensitive to the type of model defects that being considered.
Meanwhile, no noticeable visual features can be spotted from the example
residuals plots.

@@

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/plotsel-1}

\caption{A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50 percent of subjects evenly select other plots.}\label{fig:plotsel}
\end{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dataplotsel-1}

\caption{...}\label{fig:dataplotsel}
\end{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dadtaplotselfacet-1}

\caption{...}\label{fig:dadtaplotselfacet}
\end{figure}

\hypertarget{invesgating-interesting-lineups}{%
\section{invesgating interesting
lineups}\label{invesgating-interesting-lineups}}

@@

>>>>

\title{Why is it that statistical tests for residuals are not widely
used? An explanation using visual inference.}

@@

\citet{vanderplas2021statistical} suggest that \(\alpha\) can be
estimated using MLE with Equation \ref{eq:c-interesting-expectation}.
But for precise estimate of \(\alpha\), additional responses to
Rorschach lineups, which is a type of lineup that consists of

plots constructed from the same null data generating mechnisam.

@@

(handbook of mathathetical function,
\url{https://aapt.scitation.org/doi/10.1119/1.15378}, cite the package)

(hermite poly is created by \ldots{} in 18..,)

@@

As mentioned in Section \ref{calculating-p-values}, \(\alpha\) used in
Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using
Rorschach lineups. Hence, 36 lineups consist of only null plots with all
combinations of \(n\) and \(X_{raw}\) and three replications are
designed to be included in experiment III. Note that we generated the
lineups using the same simulation mechanism discussed in Section
\ref{sampling-from-the-null-distribution} rather than from a null model.

(it has zero effect size, an alternative to Rorschach lineups)

(Appendix: vary the alpha a little bit and see how the curves change,
sensitivity analysis)

Thus, we used another valid method suggested in
vanderplas2021statistical which was designed for typical lineups to
estimate \(\alpha\) correctly. However, by not using Rorschach lineups,
we did lose some accuracy in the estimation. (questionable)

We recommend future studies to generate Rorschach lineups strictly
following the definition given by buja\_statistical\_2009.

@@

(discuss choice of tests, mention pkg \texttt{lmtest},
\texttt{skedastic})

@@

For the non-linearity model, the power of all four tests increases as
the effect gets larger. The power curve of RESET test climbs
aggressively from 13\% to around 70\% as \(log_e(E)\) increases from 0
to 2, while power of other tests respond inactively to the change of
effect and remain lower than 25\% throughout the period, showing that
F-test is way more sensitive to the type of model defects that being
considered. Meanwhile, no noticeable visual features can be spotted from
the example residuals plots.

@@

@@

\hypertarget{appendix-contains-all-the-details}{%
\section{appendix (contains all the
details)}\label{appendix-contains-all-the-details}}

\hypertarget{a-collection-interesting-lineups-unusual-results}{%
\subsubsection{A collection interesting lineups (unusual
results)}\label{a-collection-interesting-lineups-unusual-results}}

why unusual? what is the possible explanations?

\hypertarget{targe-journal}{%
\section{targe journal}\label{targe-journal}}

JRSSB: Journal of the Royal Statistical Society Series B (Statistical
Methodology) Deadline: Jan 1, Apr 1

Reading: style of writing, author guideline
(\url{https://rss.onlinelibrary.wiley.com/hub/journal/14679868/author-guidelines})

JCGS JDSS
